---
title: "Introduction to the fars Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the fars Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

The `fars` package provides a small set of functions for working with data
from the US National Highway Traffic Safety Administration's Fatality
Analysis Reporting System (FARS). FARS is a nationwide census providing
yearly data on fatal injuries suffered in motor vehicle crashes.

The package includes functions to:

- read a single year's FARS data file (`fars_read`)
- build the standard FARS filename for a given year (`make_filename`)
- read and combine multiple years of FARS data (`fars_read_years`)
- summarize accident counts by month and year (`fars_summarize_years`)
- plot accident locations for a given state and year on a map
  (`fars_map_state`)

## Reading a single year of data

`fars_read()` reads a single FARS CSV (optionally `.bz2`-compressed) file
into a tibble. It expects the file to already exist locally.

```{r}
library(fars)
data_2013 <- fars_read("accident_2013.csv.bz2")
head(data_2013)
```

## Building filenames

`make_filename()` constructs the expected FARS filename for a given year,
following the `"accident_<year>.csv.bz2"` naming convention.

```{r}
make_filename(2013)
#> [1] "accident_2013.csv.bz2"
```

## Reading multiple years

`fars_read_years()` reads several years of data at once, returning a list
of tibbles (one per year) containing just the `MONTH` and `year` columns.
Years that fail to load produce a warning and a `NULL` entry rather than
stopping execution.

```{r}
fars_read_years(c(2013, 2014, 2015))
```

## Summarizing accidents by month and year

`fars_summarize_years()` builds on `fars_read_years()` to produce a wide
summary table with one row per month and one column per year, showing the
number of accidents recorded.

```{r}
fars_summarize_years(c(2013, 2014, 2015))
```

## Mapping accidents for a state

`fars_map_state()` reads a single year of data, filters to a given state
(using the FARS numeric state code), and plots the location of each
accident on a map of that state.

```{r}
# Plot accidents in California (state code 6) for 2013
fars_map_state(6, 2013)
```

## Session info

```{r, eval = TRUE}
sessionInfo()
```
